Climate indicators

### 2023-02-23 DRAFT_climate_indicators.Rmd
normalised across all columns

@@

The climate variable is the same as y1, y2, y~*n*~, ....
But to create some differences between the variable and the reference values, I will tweak these slightly, more or less divide them by two.

```{r}
#y1v <- y1 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom)
#y2v <- y2 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
#y3v <- y3 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
#y4v <- y4 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
#y5v <- y5 %>% filter(band == 1) %>% adrop() %>% st_apply(1:2, FUN = myRandom) 
#y_v <- c(y1v, y2v, y3v, y4v, y5v, y_ref)

y1v <- y1 %>% adrop() %>% mutate(Attribute_A = Attribute_A/2)
y_v <- c(y1v, select(y_ref, reference_value))
y_v_merged <- merge(y_v)
dim(y_v_merged)
names(y_v_merged)
y_v_merged <- setNames(y_v_merged, "value")

st_get_dimension_values(y_v_merged, "attributes")
```

#### Normalise variable 
I now want to to cell-by-cell computations across the _attributes_ dimension, and divide _Attribute_A_ by _reference_value_.

@@

```{r, fig.cap"The normalised indicator for year 1970"}
tm_shape(temp)+
  tm_raster()+
  tm_layout(legend.outside = T)
```
This looks like perfect noise, and it is, because the variable and reference value are so well correlated.
If the reference value was static, it would look different.

@@

```{r}
tm_shape(temp)+
  tm_raster()+
  tm_layout(legend.outside = T)
```

>>>>

#### Export
I should export the reference data at this point (write to file) as we don't need to recalculate this every time we want to update the indicator values.

@@

The climate variable is the same as y1, y2, y~*n*~, ....
But to create some differences between the variable and the reference values, I will tweak these slightly, more or less divide them by two. I then merge with the reference value and sd, creating one common data cube.

```{r}
y_var <- y %>% mutate(Attribute_A = Attribute_A/2) %>% 
  split(3) %>% c(y_ref)
y_var <- y_var %>%
  setNames(paste0("v_",substr(names(y_var), 1,4)))%>%
  rename("reference" = v_refe,
         "sd" = v_sd)
```

```{r}
# select the columns to normalise
cols <- names(y_var)[!names(y_var) %in% c("reference", "sd") ]
cols_new <- cols
names(cols_new) <- gsub("v_", "i_", cols)
# Mutate 
y_var_norm <- y_var %>%
  mutate(across(all_of(cols), ~.x / reference)) %>%
  rename(all_of(cols_new))

#y_v <- c(y1v, select(y_ref, reference_value))
#y_v_merged <- merge(y_v)
#dim(y_v_merged)
#names(y_v_merged)
#y_v_merged <- setNames(y_v_merged, "value")
#st_get_dimension_values(y_v_merged, "attributes")
```

```{r}
ggarrange(
ggplot() + 
  geom_stars(data = y_var["v_1970"]) +
  coord_equal() +
  theme_void() +
  scale_fill_viridis_c(option = "D") +  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))
,
ggplot() + 
  geom_stars(data = y_var_norm["reference"]) +
  coord_equal() +
  theme_void() +
  scale_fill_viridis_c(option = "D") +  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))
,
ggplot() + 
  geom_stars(data = y_var_norm["i_1970"]) +
  coord_equal() +
  theme_void() +
  scale_fill_viridis_c(option = "D") +  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0))
, ncol=3
)
```

#### Normalise variable 
I now want to do cell-by-cell computations across the _attributes_ dimension, and divide _Attribute_A_ by _reference_value_.

@@

```
This operation is quite fast!

```{r, fig.cap = "The normalised indicator for year 1970."}
tm_shape(temp)+
  tm_raster()+
  tm_layout(legend.outside = T)
```
This looks like perfect noise, and it is, because the variable and reference value are so well correlated.
If the reference value was static, it would look different.

@@

```{r, fig.cap = "The normalised indicator for year 1970 when the reference value is the same all over."}
tm_shape(temp)+
  tm_raster()+
  tm_layout(legend.outside = T)
```
